State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-06-10 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
paper edits and added PSD figure

where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface. The connectivity kernel is typically a ``Mexican hat'' function which describes strong local activation, weak mid-range repression and weak long-range activation. The exact shape of this kernel is assumed to vary across patients, and hence needs to be inferred from data. \ken{Risky comment if we don't get better estimates.}

@@

where $T_s$ is the time step, $\xi = 1-\zeta T_s$ and $e_t(\mathbf{r})$ is an i.i.d. disturbance such that $e_t\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$. Here $\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a spatial Gaussian Process with mean zero and covariance function $\gamma(\mathbf{r}-\mathbf{r}')$~\cite{Rasmussen2005}. This term is added to account for uncertainty in the model. To simplify the notation, the index of the future time sample, $t+T_s$, shall be referred to as $t+1$ throughout the rest of the paper. The mapping between the membrane voltage and the collected iEEG/LFP data is modelled using the observation function 

@@

where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. The choice of Gaussian basis functions can be justified by the existence of the so called bump solutions for this class of model, which have a Gaussian shape~\cite{Coombes2005}. The width and positioning of the basis functions can be determined by spectral analysis (explained in detail in Section~\ref{SpectralAnalysisSection}). The connectivity kernel can also be decomposed as 

@@

	 \lefteqn{\mathbf{x}_{t+1} \nonumber = T_s\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r}) \int_\Omega {\boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta}} \nonumber\\
&+&\xi\mathbf{x}_t + \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}(\mathbf{r})e_t(\mathbf{r})d\mathbf{r}}.

@@

\begin{figure*}
\subfigure[][]{\includegraphics{./Graph/KernelEstimate.pdf}}
 \subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT.pdf}}\\
\caption{Results for connectivity estimate. In all the subplots the true kernel is shown by the solid line, the mean kernel estimate (over 100 realisations) is shown by the dashed line and the 95~\% confidence interval is shown by the shaded grey region. \textbf{a.} Comparison of kernel estimates to true kernel in the spatial domain. \textbf{b.} Comparison of the true kernel to the estimated kernel in the frequency domain. \dean{Note: needs to be changed to absolute values if we include it.} \dean{Note: Power and dB needs to be non-italicised.}}
\label{fig:KernelEstimates}
\end{figure*}
\begin{figure}[th]
\subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the connectivity kernel basis function parameter ratios over 100 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20).}
\end{figure}
\label{fig:ParametersRatio}
\begin{figure}[th]
\centering
\subfigure[][]{\label{fig:FieldFFT}\includegraphics[scale=1]{./Graph/FFTTrueField.pdf}}
\subfigure[][]{\label{fig:EstimatedFieldFFT}\includegraphics[scale=1]{./Graph/FFTEstimatedField.pdf}}\\
\caption{Spatial frequency of the neural fields. \textbf{a.} The average (over time) spatial frequency of the neural field. \textbf{b.} The average (over time) spatial frequency of the estimated neural field resulting from the basis function decomposition. As a result of the band limiting properties of the sensors and basis functions the reconstructed shows a narrower bandwidth.}
\label{fig:FFTTrueEstimate}
\end{figure}

>>>>

where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface. The connectivity kernel is typically a ``Mexican hat'' function which describes strong local activation, weak mid-range repression and weak long-range activation. The exact shape of this kernel is assumed to vary across patients, and hence needs to be inferred from data.

@@

where $T_s$ is the time step, $\xi = 1-\zeta T_s$ and $e_t(\mathbf{r})$ is an i.i.d. disturbance such that $e_t\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$. Here $\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a spatial Gaussian process with mean zero and covariance function $\gamma(\mathbf{r}-\mathbf{r}')$~\cite{Rasmussen2005}. This term is added to account for uncertainty in the model. To simplify the notation, the index of the future time sample, $t+T_s$, shall be referred to as $t+1$ throughout the rest of the paper. The mapping between the membrane voltage and the collected iEEG/LFP data is modelled using the observation function 

@@

where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. %The choice of Gaussian basis functions can be justified by the existence of the so called bump solutions for this class of model, which have a Gaussian shape~\cite{Coombes2005}. 
The width and positioning of the basis functions can be determined by spectral analysis (explained in detail in Section~\ref{SpectralAnalysisSection}). The connectivity kernel can also be decomposed as 

@@

	 \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{-1}\int_\Omega \boldsymbol{\phi}(\mathbf{r}) \int_\Omega \boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \nonumber \\ 
	\times d\mathbf{r}' d\mathbf{r} \boldsymbol{\theta} + \xi\mathbf{x}_t + \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}(\mathbf{r})e_t(\mathbf{r})d\mathbf{r}}.

@@

\begin{figure*}
\subfigure[][]{\includegraphics{./Graph/KernelEstimate.pdf}}
 % \subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT.pdf}}\\
\caption{Results for connectivity estimate. In all the subplots the true kernel is shown by the solid line, the mean kernel estimate (over 100 realisations) is shown by the dashed line and the 95~\% confidence interval is shown by the shaded grey region. \textbf{a.} Comparison of kernel estimates to true kernel in the spatial domain. \textbf{b.} Comparison of the true kernel to the estimated kernel in the frequency domain. \dean{Note: needs to be changed to absolute values if we include it.} \dean{Note: Power and dB needs to be non-italicised.}}
\label{fig:KernelEstimates}
\end{figure*}
\begin{figure}[th]
\subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the connectivity kernel basis function parameter ratios over 100 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20).}
\end{figure}
\label{fig:ParametersRatio}
\begin{figure}[th]
\centering
\subfigure[][]{\label{fig:FieldFFT}\includegraphics[scale=1]{./Graph/FFTTrueField.pdf}}
\subfigure[][]{\label{fig:EstimatedFieldFFT}\includegraphics[scale=1]{./Graph/FFTEstimatedField.pdf}}\\
\caption{Spatial frequency of the neural fields. \textbf{a.} The average (over time) spatial frequency of the neural field. \textbf{b.} The average (over time) spatial frequency of the estimated neural field resulting from the basis function decomposition. As a result of the band limiting properties of the sensors and basis functions the reconstructed shows a narrower bandwidth.}
\label{fig:FFTTrueEstimate}
\end{figure}
